Binary companions of red giant stars in \apogee\ \DR

### 2018-02-27 paper/1-catalog/twoface-catalog.tex
Move some text

We perform our individual system fits---that is, make our posterior
samplings---under the following assumptions:
\begin{description}
\item[no multiplets] All radial-velocity variations of the primary star are
  induced by a single companion.
  This is motivated by the idea that triple star systems are usually
  hierarchical so that the period of the inner binary is typically much shorter
  than the orbital period of the outer body (\todo{CITATION NEEDED}).
  At present, we ignore the possibility of higher-order multiple systems.
  % \todo{revisit this - do we want to also sample over a long-term trend?}
\item[Kepler] Related to the first point, all velocity variations of the primary
  are gravitational, and we therefore ignore the possibility of coherent intrinsic variation from, e.g., stellar oscillations.
\item[SB1] All spectra are single-lined; that is, we assume that the secondary
  is significantly fainter and is thus undetected in the spectra.
  This assumption is motivated by the fact that we expect \RGB\ stars to be
  substantially more luminous than their typical companion.
  However, there are known main-sequence double-lined binary stars in the
  \apogee\ catalog \citealt{El-Badry:2018}, and an expected but unknown fraction
  of \RGB--\RGB\ binaries.
\item[simple noise model] Measurements are unbiased and noise estimates are
  correct up to an unknown excess variance.
  All noise contributions result in Gaussian uncertainties on the individual
  radial-velocity measurements.
\end{description}

@@

\subsection{Implementation notes}

\todo{Nontrivial computation - details about running on cluster, TheJoker vs. emcee}

@@

Using the maximum-likelihood hyperparameters derived from the initial posterior
samplings (see \sectionname~\ref{sec:inferjitter}), we update and fix the
excess variance prior distribution parameters---$(\mu_{y}, \sigma_{y}) = (9.50,
1.64)$, in units of $\textrm{m}~\textrm{s}^{-1}$---and re-run \thejoker\ on the
parent sample of \apogee\ \DR\ stars.
\todo{Discuss how this relates to empirical Bayes?}

From this run, 91,096 stars completed and successfully returned \nposterior\
samples using \thejoker\ alone.
The remaining 5,135 stars did not return \nposterior\ samples: 4,744 were
flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}; We
continue generating samples for these stars using the methodology explained
above (see \sectionname~\ref{sec:samplings}).
In the end, \todo{XX} stars have completed samplings with \nposterior\ samples,
and, after re-running \thejoker\ with twice as many prior samples, \todo{XX}
stars still have fewer samples that are multi-modal in period.

@@

\subsection{Stars with confident companions}
\label{sec:catalog-confident}

We don't expect there to be a sharp transition in orbital parameters between
stars with companions and stars without companions: There is a continuum of
companion properties.
For example, the companion masses can be brown dwarf-like, or at long periods,
or at high inclination, all of which will make the velocity semi-amplitude, $K$,
small for a given system.
From the radial velocity data alone, there is therefore no way to select a
complete sample of stars with and without companions.

\todo{Thresholding on what now?}

@@

>>>>

We perform our individual system fits---that is, make our posterior
samplings---under the following assumptions:
\begin{description}
\item[no multiplets] All radial-velocity variations of the primary star are
  induced by a single companion.
  This is motivated by the idea that triple star systems are usually
  hierarchical so that the period of the inner binary is typically much shorter
  than the orbital period of the outer body (\todo{CITATION NEEDED}).
  At present, we ignore the possibility of higher-order multiple systems.
  % \todo{revisit this - do we want to also sample over a long-term trend?}
\item[Kepler] Related to the first point, all velocity variations of the primary
  are gravitational, and we therefore ignore the possibility of coherent intrinsic variation from, e.g., stellar oscillations.
\item[SB1] All spectra are single-lined; that is, we assume that the secondary
  is significantly fainter and is thus undetected in the spectra.
  This assumption is motivated by the fact that we expect \RGB\ stars to be
  substantially more luminous than their typical companion.
  However, there are known main-sequence double-lined binary stars in the
  \apogee\ catalog (\citealt{El-Badry:2018}), and an expected but unknown
  fraction of \RGB--\RGB\ binaries.
\item[simple noise model] Measurements are unbiased and noise estimates are
  correct up to an unknown excess variance.
  All noise contributions result in Gaussian uncertainties on the individual
  radial-velocity measurements.
\end{description}

@@

\subsection{Control sample}
\label{sec:control-sample}

\todo{describe how we make it}

% \subsection{Implementation notes}
%
% \todo{Nontrivial computation - details about running on cluster, TheJoker vs. emcee}

@@

Using the maximum-likelihood hyperparameters derived from the initial posterior
samplings (see \sectionname~\ref{sec:inferjitter}), we update and fix the
excess variance prior distribution parameters---$(\mu_{y}, \sigma_{y}) = (9.50,
1.64)$, in units of $\textrm{m}~\textrm{s}^{-1}$---and re-run \thejoker\ on the
parent sample of \apogee\ \DR\ stars.
This methodology, in which we estimate a prior from the data and then fix it, is
typically referred to as ``empirical Bayes''; In this case, it is an
approximation to doing a hierarchical inference of the individual system orbital
parameters and the excess variance distribution hyperparameters simultaneously.

From this run, 91,096 stars completed and successfully returned \nposterior\
samples using \thejoker\ alone.
The remaining 5,135 stars did not return \nposterior\ samples: 4,744 were
flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}.
We continue generating samples for these stars using the methodology explained
above (see \sectionname~\ref{sec:samplings}).
In the end, \todo{XX} stars have completed samplings with \nposterior\ samples,
and, after re-running \thejoker\ with twice as many prior samples, \todo{XX}
stars still have fewer samples that are multi-modal in period.

@@

\subsection{Stars with companions}
\label{sec:catalog-confident}

We don't expect there to be a sharp transition in orbital parameters between
stars with companions and stars without companions: There is a continuum of
companion properties.
For example, the companion masses can be low mass, or at long periods, or at
high inclination, all of which will make the velocity semi-amplitude, $K$, small
for a given system.
From the radial velocity data alone, there is therefore no way to define a
simple cut to select a complete sample of stars with companions.
Nevertheless, it is possible to define a cut that selects stars with
high-confidence companions while similarly minimizing an estimated
false-positive rate.

\todo{False-positive estimate using the null control sample.}

We select a sample of stars that confidently have companions using percentiles
computed from the posterior samples in the velocity semi-amplitude parameter,
$K$.
\figurename~\ref{fig:} shows distributions of percentiles in $\ln K$ computed
for all stars with posterior samplings from \thejoker.
\todo{percentiles figure}

@@

\todo{\emph{no multiplets} punt to future work}

\todo{\emph{Kepler} evidence for coherent oscillations at the upper giant branch? companions that don't make sense}

\todo{\emph{SB1} comment that there are some companions with masses large enough that they should be evolved, and therefore should contribute to the light! where are the companions? neutron stars / BHs?}
